---
title: "Bird_Speed_Table"
author: "Daina Bouquin"
date: "November 30, 2015"
output: html_document
---

The RPubs version of the markdown file is available [here](http://rpubs.com/dbouquin/IS607_final_birdspeedtable)

```{r, warning=FALSE, message=FALSE}
library(rvest)
library(stringr)
library(knitr)
```

These scripts can be used to scrape a table by copying the table's xpath in the source. 
I obtained this xpath using Chrome's developer tools
```{r, warning=FALSE}
url <- "https://en.wikipedia.org/wiki/List_of_birds_by_flight_speed"

bird_speeds <- url %>%
  html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/table[1]') %>%
  html_table()
bird_speeds <- bird_speeds[[1]]

colnames(bird_speeds)
# Rename the columns
colnames(bird_speeds) <- c("bird", "image", "species", "family", "avg_h_speed", "max_h_speed", "avg_d_speed", "max_airspeed", "flight")

head(bird_speeds)
bird_speeds[1,]
```

To clean up the speed columns, I'll remove the citation tags in square brackets using gsub
```{r}
bird_speeds$avg_h_speed <- gsub("\\[(.+)\\]", "", bird_speeds$avg_h_speed)
bird_speeds$max_h_speed <- gsub("\\[(.+)\\]", "", bird_speeds$max_h_speed)
bird_speeds$avg_d_speed <- gsub("\\[(.+)\\]", "", bird_speeds$avg_d_speed)
bird_speeds$max_airspeed <- gsub("\\[(.+)\\]", "", bird_speeds$max_airspeed)
head(bird_speeds)
```

I will also now split up the speed columns so that there are separate columns for mph and km and the units aren't embedded in the data

```{r}
# average horizontal speed
bird_speeds$avg_h_speed_mph <- lapply(strsplit(as.character(bird_speeds$avg_h_speed), "km/h"), "[", 2)
bird_speeds$avg_h_speed_km <- lapply(strsplit(as.character(bird_speeds$avg_h_speed), "km/h"), "[", 1)
head(bird_speeds)
# Remove the units and the new line characters
bird_speeds$avg_h_speed_mph <- gsub("\n", "", bird_speeds$avg_h_speed_mph)
bird_speeds$avg_h_speed_mph <- gsub("mph", "", bird_speeds$avg_h_speed_mph)
# remove the old avg_h_speed column and others likely to not be useful
bird_speeds <- subset(bird_speeds, select = c(bird,species,family,max_h_speed,avg_d_speed,max_airspeed,avg_h_speed_mph, avg_h_speed_km))

# Check the structure
head(bird_speeds)
```
Perform these tasks again with other speed measure columns
```{r}
# max horizontal speed
bird_speeds$max_h_speed_mph <- lapply(strsplit(as.character(bird_speeds$max_h_speed), "km/h"), "[", 2)
bird_speeds$max_h_speed_km <- lapply(strsplit(as.character(bird_speeds$max_h_speed), "km/h"), "[", 1)
bird_speeds$max_h_speed_mph <- gsub("\n", "", bird_speeds$max_h_speed_mph)
bird_speeds$max_h_speed_mph <- gsub("mph", "", bird_speeds$max_h_speed_mph)

# average dive speed
bird_speeds$avg_d_speed_mph <- lapply(strsplit(as.character(bird_speeds$avg_d_speed), "km/h"), "[", 2)
bird_speeds$avg_d_speed_km <- lapply(strsplit(as.character(bird_speeds$avg_d_speed), "km/h"), "[", 1)
# Remove the units and the new line characters
bird_speeds$avg_d_speed_mph <- gsub("\n", "", bird_speeds$avg_d_speed_mph)
bird_speeds$avg_d_speed_mph <- gsub("mph", "", bird_speeds$avg_d_speed_mph)

# max airspeed
bird_speeds$max_airspeed_mph <- lapply(strsplit(as.character(bird_speeds$max_airspeed), "km/h"), "[", 2)
bird_speeds$max_airspeed_km <- lapply(strsplit(as.character(bird_speeds$max_airspeed), "km/h"), "[", 1)
# Remove the units and the new line characters
bird_speeds$max_airspeed_mph <- gsub("\n", "", bird_speeds$max_airspeed_mph)
bird_speeds$max_airspeed_mph <- gsub("mph", "", bird_speeds$max_airspeed_mph)
head(bird_speeds)

# Select only the columns we want
bird_speeds <- subset(bird_speeds, select = c(bird,species,family,avg_h_speed_mph, avg_h_speed_km,max_h_speed_mph, max_h_speed_km, avg_d_speed_mph, avg_d_speed_km, max_airspeed_mph, max_airspeed_km ))
head(bird_speeds)
```
There also appear to be citation brackets in the family column so I'll remove them
```{r}
bird_speeds$family <- gsub("\\[(.+)\\]", "", bird_speeds$family)
```
Now I'll show the resulting dataframe
```{r}
kable(head(bird_speeds), format = "markdown")
```
