---
title: "Pigeons Data Mining and Cleaning"
author: "Nabila Hossain"
date: "December 13, 2015"
output: html_document
---

###_The following are the steps that I took, into to creating the .CSV file for the Dickin Medal winning pigeons and combinning the racing pigeons dataset._   
  
* ####[Uploading the Dickin Medal Winning animals data from Wikipedia, and Creating a CSV with only the information about the pigeons.](#1)  
* ####[Mannually inputing the other infromation](#2)  
* ####[Extract the three CSV's from GitHub, and commbinig them.](#3)  
* ####[Tiding and cleaning the data and creaing a new CSV.](#4)  


<a name="1"/>__The PDSA Dickin Medal was instituted in 1943 in the United Kingdom by Maria Dickin to honor the work of animals in World War II. The award is commonly referred to as "the animals' Victoria Cross".The medal was awarded to 32 pigeons, 18 dogs, three horses, and one cat, for their outstanding service during WWII. Using `rvest`, `stringr`, and `dplyr` packages I uploaded the Wikipedia's table about the recipients of Dickin Medal. I created a CSV with only the pigeons for this assignment.__

```{r}
library(rvest)
library(stringr)
library(dplyr)

url <- read_html("https://en.wikipedia.org/wiki/Dickin_Medal", Encoding = "UTF-8")
table1 <- url %>% html_nodes("table") %>% .[[2]]  %>% html_table()
head(table1, 4)
names(table1) <- c("Recipiant", "Animal", "Date_of_award", "Ref")

table1$Date_of_award <- str_replace_all(table1$Date_of_award, pattern = "\\d{8,}.*?([1-9] \\w{3,}) (\\d{4})", replacement = "\\1 \\2") 
table1$Date_of_award <- str_replace_all(table1$Date_of_award, pattern = "\\d{8,}.*?\\d{3,}(\\w{3,} \\d{4})", replacement = "\\1")
table1$Date_of_award <- str_replace_all(table1$Date_of_award, pattern = "\\d{8,}.*?\\-[0]{4}(\\d{4})", replacement = "\\1")

table2 <- table1 %>% subset(Animal == "Pigeon") %>% select(Recipiant, Animal, Date_of_award)
  
head(table2, 10)

#write.csv(table2, file = "C:/Users/Nabila/Documents/CLASS  IS 607/Final/Dickin_Medal_Pigeons_original.csv")
```


<a name="2"/>__After creating the CSV I manually obtained more information from various sources about Dickin Medal Pigeon Winners, including further information about their service, distance they traveled and the time they took to accomplish their assignment. Below is a preview of the final CSV.__

```{r}
DMT <- read.csv(file="https://raw.githubusercontent.com/dbouquin/Pigeons/master/Dickin_Medal_Pigeons.csv", header=TRUE, sep=",")
head(DMT)
```

<a name="3"/>__I uploaded the three CSV files from our GitHub repository, which holds the data for the racing pigeons today (2011-2015). Then I tidied the CSV files so that I can combine them using `rbind`.__

```{r}
library(dplyr)
library(stringr)
library(car)

YB_url <- read.csv(file="https://raw.githubusercontent.com/dbouquin/Pigeons/master/American_Racing_Pigeon_Union_Young_Bird_Data.csv", header=TRUE, sep=",")
OB_url <- read.csv(file="https://raw.githubusercontent.com/dbouquin/Pigeons/master/OldBirds.csv", header=TRUE, sep=",")
YB2_url <- read.csv(file="https://raw.githubusercontent.com/dbouquin/Pigeons/master/PIGEONDATAYBRM.csv", header=TRUE, sep=",")


YB1_table <- data.frame(YB_url[c(5, 8:9, 11, 13)])
YB1_table$TYPE <- "YOUNG BIRD"
OB1_table <- data.frame(OB_url[c(4, 9:10, 12, 14)])
OB1_table$TYPE <- "OLD BIRD"
YB2_table <- data.frame(YB2_url[c(3, 6:7, 10, 12)])
colnames(YB2_table) <- c("NAME", "CLR", "X", "MILES", "YPM")
YB2_table$TYPE <- "YOUNG BIRD"
YB2_table$MILES <- as.factor(YB2_table$MILES)

head(YB1_table)
head(OB1_table)
head(YB2_table)

Bird_table <- rbind.data.frame(YB1_table, OB1_table, YB2_table)
nrow(Bird_table)
```

<a name="4"/>__Using `car` package I combined and changed the color and sex columns. Then I created a CSV file and uploaded it to GitHub.__
```{r}
Bird_table$CLR <- str_replace_all(Bird_table$CLR, "(\\_{1,})|(\\s{1,})|(\\`)", replacement = "")
Bird_table$YPM <- str_replace_all(Bird_table$YPM, "(\\s{1,})", replacement = "")
Bird_table$NAME <- str_replace_all(Bird_table$NAME, "(\\/[[:digit:]])|(\\s{1,}\\/[[:digit:]])", replacement = "")

head(Bird_table, 15)

Bird_table$CLR <- as.factor(Bird_table$CLR)
Bird_table$CLR <- recode(Bird_table$CLR, "c('B', 'BB', 'BBAR', 'BBPD', 'BBSP', 'BBTC', 'BBWF', 'DBB')='blue bar'; c('BC', 'BCSP', 'BCPD', 'BCWF')='blue check'; c('DC', 'DCSP', 'DCPD', 'DCWF')='dark check'; c('RB', 'RBSP', 'RBPD', 'RBWF')='red bar'; c('RC', 'RCSP', 'RCPD', 'RCWF')='red check'; c('RED', 'RSPL')='red'; c('SPL','BSPL', 'RSPL')='splash'; c('BLK')='black'; c('WHT')='white'; c('GRIZ')= 'grizzled'; c('IND')='indigo'; c('LAV')='lavender'; c('AND')='andalusian'; c('SLT')='slate'; c('BRWN', 'CHOC', 'BRNZ')='brown var'; c('TIC')= 'ticked'; c('PEN', 'YLW')= 'yellow var'; c('', '1', 'UNKN')='unknown'; else='complex'")
Bird_table$X <- recode(Bird_table$X, "'H'= 'hen'; 'C'='cock'; else='unknown'")


Bird_table$YPM <- round(as.numeric(Bird_table$YPM), 1)
colnames(Bird_table) <- c("TEAM_NAME", "COLOR", "SEX", "DISTANCE_MILES", "SPEED_YPM", "TYPE")

head(Bird_table, 15)

#write.csv(Bird_table, file = "C:/Users/Nabila/Documents/GitHub/Pigeons/Combined_Racing_Pigeons_Table.csv")
```

