---
title: "fast_animals_table"
author: "Daina Bouquin"
date: "December 5, 2015"
output: html_document
---

The rendered R Markdown for this file can be viewed [here](http://rpubs.com/dbouquin/IS607_final_fastanimals_table).
    
The results of this data cleaning can be found on GitHub [here](https://github.com/dbouquin/Pigeons/blob/master/fast_animals.csv). The original data source is a Wikipedia page [here](https://en.wikipedia.org/wiki/Fastest_animals)-- this code was applied to the first table in the article.

```{r, warning=FALSE, message=FALSE}
library(rvest)
library(dplyr)
library(stringr)
```
```{r}
url <- "https://en.wikipedia.org/wiki/Fastest_animals"
fast_animals <- url %>%
  read_html() %>%
  html_nodes(xpath='//*[@id="mw-content-text"]/table[1]') %>%
  html_table(fill = TRUE)
fast_animals <- fast_animals[[1]]
```
Check out the structure and start cleaning up the table
```{r}
head(fast_animals)
str(fast_animals)
colnames(fast_animals)
# drop the columns we do not need
drops <- c("Sl.no", "Notes")
fast_animals <- fast_animals[,!(names(fast_animals) %in% drops)]
str(fast_animals)
head(fast_animals)
# rename columns
colnames(fast_animals) <- c("animal", "max_speed", "class")
```
All we need is the mph data, so we'll extract it out from the () in the Speed column and create a new column with just this value
```{r}
# remove all [number] pattern from the animal column
fast_animals$animal <- gsub("\\[(.+)\\].+?", "", fast_animals$animal)
fast_animals$animal <- gsub("\\[(.+)\\]", "", fast_animals$animal)
# remove all but parens in the speed column
fast_animals$max_speed <- gsub("[\\(\\)]", "", regmatches(fast_animals$max_speed, gregexpr("\\(.*?\\)", fast_animals$max_speed)))
# Remove mph from speed column and take just the highest max speed
fast_animals$max_speed <- gsub("mph", "", fast_animals$max_speed)
# Remove the white space
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
fast_animals$max_speed <- trim(fast_animals$max_speed)
# manually clean up two range records
fast_animals[2,2] <- (150+200)/2
fast_animals[12,2] <- (68+75)/2
fast_animals$max_speed <- as.numeric(fast_animals$max_speed)
fast_animals
```
Now I will convert the speed column to yards/minute and write the results to a CSV file.
```{r}
fast_animals <- mutate(fast_animals, max_speed_ym = ((max_speed*1760)/60))
# Rename/reorder the columns
colnames(fast_animals) <- c("animal", "max_speed_mph", "class", "max_speed_ym")
fast_animals <- select(fast_animals, animal, class, max_speed_mph, max_speed_ym) #reorder
fast_animals

# write restults to CSV
write.csv(fast_animals, "fast_animals.csv")
```